---
title: "Fitting N-of-1 models to data"
author: "Ed Baskerville"
date: "4/25/2018"
output:
  html_document:
    fig_width: 6
    fig_height: 2
---

This file demonstrates how to fit models to N-of-1 clinical trial simulations.
Note that none of these are explicit time-series process models, which is a major limitation.

```{r}
source('../n1-simulator.R')
library(ggplot2)
```

First, we'll generate the same time series as in `1-simulation.Rmd`:

```{r}
result <- n1_simulate(
  n_treatments = 2, n_blocks = 2,
  baseline_initial = 160, effect_size = c(-40, -30), tc_in = c(3, 2), tc_out = c(2, 3), tc_outcome = 1,
  sd_baseline = 0.4, sd_outcome = 0.6, sd_obs = 4.0,
  treatment_period = 30.0, sampling_timestep = 1.0, noise_timestep = 0.01,
  treatment_order = NULL, random_seed = 1, return_data_frame = TRUE
)
timeseries <- result$timeseries
```

```{r}
ggplot(data = timeseries, aes(x = t)) + geom_line(aes(y = outcome)) + geom_point(aes(y = outcome_obs))
```

## Model 1: treatment only

The simplest AHRQ model regresses the outcome

```{r}
lm(outcome_obs ~ factor(treatment), data = timeseries)
```

The intercept represents the baseline plus effect of treatment 1, and the coefficient for treatment 2 represents the effect size difference between the two treatments.

This is encapsulated in the function `n1_model1`, which returns a named list containing three items:
* `coefficients`: a data frame with effect coefficient estimates in the `estimate` column, and p-values in the `pvalue` column
* `fit_obj`: the raw object returned by `lm`
* `AIC` the Akaike Information Criterion value (AIC) for the model, for selecting among models

```{r}
n1_fit_model1(n_treatments = 2, timeseries)
```

## Model 2: treatment + block

Model 2 adds a per-block regression term:

```{r}
lm(outcome_obs ~ factor(treatment) + factor(block), data = timeseries)
```

A shortcut is provided as `n1_model2`:

```{r}
n1_fit_model2(n_treatments = 2, timeseries)
```

Note the lower AIC value for this model, indicating that baseline drift is detected.

## Model 3: treatment + block + treatment * time

Model 3 adds an interaction with time:

```{r}
lm(outcome_obs ~ factor(treatment) + t * factor(treatment) + factor(block), data = timeseries)
```

and gives analogous results with `n1_fit_model3`:

```{r}
n1_fit_model3(n_treatments = 2, timeseries)
```

Again, note that the AIC value shows an improvement over the simpler models.


## Model 4: treatment + block mixed effect

Model 4 is a mixed-effect model, and requires the `nlme` package:

```{r}
library(nlme)
lme(fixed = (outcome_obs ~ factor(treatment)), random = (~ 1 | block), data = timeseries, method = 'ML')
```

and it is accessible via `n1_fit_model4`:

```{r}
n1_fit_model4(n_treatments = 2, timeseries)
```

The AIC value is bizarrely high.
