---
title: "Simulating large N-of-1 experiments"
author: "Ed Baskerville"
date: "4/25/2018"
output:
  html_document:
    fig_width: 6
    fig_height: 2
---

This file demonstrates how to simulate and fit models to many N-of-1 clinical trials at once,
sweeping over parameters and running multiple replicates per parameter set.

```{r}
source('../n1-simulator.R')
library(ggplot2)
```

## Setting up parameter values

To run a large experiment, first we use `n1_expand_parameters` to create a data frame containing parameter values.

For example, if you want to run 100 simulations for all combinations of 1 to 6 blocks and two different values for `effect_size[2]`, do this:

```{r}
parameters <- n1_expand_parameters(
  n_treatments = 2,
  n_blocks = 1:6,
  baseline_initial = 160,
  effect_size = list(
    -40, c(-30, -20)
  ),
  tc_in = c(3, 2), tc_out = c(2, 3), tc_outcome = 1,
  sd_baseline = 1.0, sd_outcome = 0.6, sd_obs = 2.0,
  treatment_period = 30.0, sampling_timestep = 1.0, noise_timestep = 0.05,
  n_replicates = 20
)
```
`n_treatments` and `n_replicates` must be held constant. All other parameters can be varied.

Vector-valued parameters must be replaced by lists in order to provide multiple values.


## Running the experiment

To simulate these parameter combinations and fit all 4 built-in models to them (see `2-fitting.Rmd`), do this:

```{r}
results <- n1_run_experiment(
  n_treatments = 2, parameters,
  initial_random_seed = 1, cores = 1, return_data_frame = TRUE
)
```

With `return_data_frame = TRUE`, results are provided in a data frame, where each row contains the results for a single clinical trial simulation.
Vector- and matrix-valued items are split into multiple columns, e.g., `effect_size_1, effect_size_2`.

With `return_data_frame = FALSE`, results are provided as a list of arrays with initial dimension equal to the number of runs, and subsequent dimensions dependent on the value (e.g., an `n_trials` X 4 X `n_treatments` array for model estimates and p-values).

Model results are given in, e.g., `estimate_1_2`, `pvalue_1_2` for model 1, coefficient 2, and `AIC_1` for model 1.

Note that the coefficient 2 corresponds to the effect size difference between treatment 1 and treatment 2.

Now, we can easily plot the experiment by the two parameters we varied and color by 0.05 significance level:

```{r}
ggplot(data = results, aes(
  x = n_blocks, y = estimate_1_2,
  color = (pvalue_1_2 < 0.05)
))+ geom_jitter() + facet_grid(. ~ (effect_size_2 - effect_size_1))
```

