---
title: "Batch mode notes"
author: "Ed Baskerville"
date: "4/24/2018"
output: html_document
---

We'll need simulation and fitting components:

```{r}
source('../n1simulator/R/simulation.R')
source('../n1simulator/R/fitting.R')
```

Each run in batch mode will require us to simulate and fit all 4 models:

```{r}
n1_simulate_and_fit <- function(
  n_blocks, n_treatments,
  baseline_initial, effect_size_vec,
  tc_in_vec, tc_out_vec, tc_outcome,
  sd_baseline, sd_outcome, sd_obs,
  treatment_period, sampling_timestep, noise_timestep,
  treatment_mat_by_block = NULL,
  random_seed = NA,
  baseline_func = NULL
) {
  data <- n1_simulate(
    n_blocks, n_treatments, baseline_initial, effect_size_vec, tc_in_vec, tc_out_vec, tc_outcome,
    sd_baseline, sd_outcome, sd_obs,
    treatment_period, sampling_timestep, noise_timestep,
    return_data_frame = TRUE, treatment_mat_by_block,
    random_seed, baseline_func
  )
  
  fits <- lapply(1:4, function(i) n1_fit(n_treatments, data, i))
  list(
    estimates = do.call(cbind, lapply(fits, function(fit) fit$coefficients[,'estimate'])),
    pvalues = do.call(cbind, lapply(fits, function(fit) fit$coefficients[,'pvalue']))
  )
}
```

Let's test it:

```{r}
result <- n1_simulate_and_fit(
  n_blocks = 2, n_treatments = 2,
  baseline_initial = 160, effect_size_vec = c(-40, -30),
  tc_in_vec = c(3, 2), tc_out_vec = c(2, 3), tc_outcome = 1,
  sd_baseline = 0.4, sd_outcome = 0.6, sd_obs = 2.0,
  treatment_period = 30.0, sampling_timestep = 1.0, noise_timestep = 0.05
)
result
```


## Generate parameters

For per-treatment parameters, we'll need this function to convert a list of vectors into arguments for `expand.grid`:

```{r}
paramlist_to_expandargs <- function(paramlist, prefix) {
  setNames(
    as.list(paramlist),
    sapply(1:length(paramlist), function(i) sprintf("%s_%d", prefix, i))
  )
}
```

Here's a function to expand parameters from vectors (and, for per-treatment parameters, lists of vectors) into a data frame:

```{r}
n1_expand_parameters <- function(
  n_treatments,
  n_blocks,
  baseline_initial, effect_size,
  tc_in, tc_out, tc_outcome,
  sd_baseline, sd_outcome, sd_obs,
  treatment_period, sampling_timestep, noise_timestep,
  n_replicates
) {
  stopifnot(length(n_treatments) == 1)
  
  stopifnot(length(effect_size) == n_treatments)
  stopifnot(length(tc_in) == n_treatments)
  stopifnot(length(tc_out) == n_treatments)
  
  # Produce all combinations of parameters
  params <- rev(do.call(
    expand.grid,
    rev(c(
      list(
        n_blocks = n_blocks,
        baseline_initial = baseline_initial
      ),
      paramlist_to_expandargs(effect_size, 'effect_size'),
      paramlist_to_expandargs(tc_in, 'tc_in'),
      paramlist_to_expandargs(tc_out, 'tc_out'),
      list(
        tc_outcome = tc_outcome,
        sd_baseline = sd_baseline,
        sd_obs = sd_obs,
        sd_outcome = sd_outcome,
        treatment_period = treatment_period,
        sampling_timestep = sampling_timestep,
        noise_timestep = noise_timestep,
        replicate_id = 1:n_replicates
      )
    ))
  ))
  params
}
```

```{r}
params <- n1_expand_parameters(
  n_treatments = 2,
  n_blocks = 1:6,
  baseline_initial = 160, effect_size = list(-40, c(-30, -20)),
  tc_in = c(3, 2), tc_out = c(2, 3), tc_outcome = 1,
  sd_baseline = 0.4, sd_outcome = 0.6, sd_obs = 2.0,
  treatment_period = 30.0, sampling_timestep = 1.0, noise_timestep = 0.05,
  n_replicates = 2
)
params
```

```{r}
get_vec_columns <- function(df, prefix, n_cols) {
  as.matrix(df[sapply(1:n_cols, function(i) sprintf('%s_%d', prefix, i))])
}
```

```{r}
n1_run_experiment <- function(n_treatments, parameters, initial_random_seed = NA, baseline_func = NULL) {
  if(is.na(initial_random_seed)) {
    initial_random_seed <- sample(2^31 - 1, 1)
  }
  
  p <- parameters
  n_trials <- nrow(p)
  trial_ids <- 1:n_trials
  random_seeds <- initial_random_seed:(initial_random_seed + n_trials - 1)
  
  effect_size_mat <- get_vec_columns(p, 'effect_size', n_treatments)
  print(effect_size_mat[1,])
  tc_in_mat <- get_vec_columns(p, 'tc_in', n_treatments)
  tc_out_mat <- get_vec_columns(p, 'tc_out', n_treatments)
  
  lapply(
    trial_ids,
    function(i) {
      n1_simulate_and_fit(
        p$n_blocks[i], n_treatments,
        p$baseline_initial[i], effect_size_mat[i,],
        tc_in_mat[i,], tc_out_mat[i,], p$tc_outcome[i],
        p$sd_baseline[i], p$sd_outcome[i], p$sd_obs[i],
        p$treatment_period[i], p$sampling_timestep[i], p$noise_timestep[i],
        treatment_mat_by_block = NULL,
        random_seed = random_seeds[i],
        baseline_func = baseline_func
      )
    }
  )
}
```


```{r}
system.time(n1_run_experiment(2, params))
```
